---
title: "Assignment 4"
author: "Kyle Ligon"
date: "`r Sys.Date()`"
output:
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
  tufte::tufte_html: default
  tufte::tufte_book:
    citation_package: natbib
    latex_engine: xelatex
link-citations: yes
---

```{r setup, include=FALSE}
library(tufte)
#invalidate cache when the tufte version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)
```

# Required Packages
```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(broom)
library(nortest)
```


# Dataset #1: Using KS on a dataset to check Exponential Distribution with Rate == 1
```{r, warning=FALSE, message=FALSE}
air <- as.tibble(datasets::airquality)
air <- filter(air, !is.na(air$Ozone))

air
```
## Hypotheses    
$H_{0}$: $F(x) = F^{*}(x)$    
$H_{1}$: $F(x) \ne F^{*}(x)$     

## Test Statistic    
```{r, warning=FALSE, message=FALSE}
ks_test <- ks.test(x = air$Ozone, 'pexp', rate = 1) %>%
  tidy()
```

We have `r round(ks_test$statistic, 4)` as our test statistic.    

## P-value    

We have `r round(ks_test$p.value, 4)` as our p-value.    

## Conclusion   
With a p-value less than 0.05, we have enough evidence to reject the null hypothesis that our data fit the exponential distribution with a Rate == 1.  There appears to be evidence that our data do not fit an exponential model with Rate == 1.  

# Dataset #2: Using the Chi-Square test to see if my data is Binomially Distributed
```{r, warning=FALSE, message=FALSE}
faithful <- as.tibble(datasets::faithful)

faithful
```

## Hypotheses   
$H_{0}: P(X$ is in class $j) = p_{j}*$    
$H_{1}: P(X$ is in class $j) \neq$ for at least one class

## Test Statistic
```{r, message=FALSE, warning=FALSE}

baseball <- c(18, 17, 16, 15, 14, 14, 13, 12, 11, 11, 10, 10, 10, 10, 10, 9, 8, 7)
hist <- hist(baseball)

breaks <- hist$breaks
counts <- hist$counts

binom_calculator <- function(left_val, right_val, n, p){
  if(right_val != 1){
    pbinom(right_val, p = p, size =  n) - pbinom(left_val, p = p, size = n)
  } else {
    1 - pbinom(left_val, p = p, size = n)
  }
}

p2 <- binom_calculator(6, 8.0, n = 45, p = 0.25)
p3 <- binom_calculator(8, 10, n = 45, p = 0.25)
p4 <- binom_calculator(10, 12, n = 45, p = 0.25)
p5 <- binom_calculator(12, 14, n = 45, p = 0.25)
p6 <- binom_calculator(14, 16, n = 45, p = 0.25)
p7 <- binom_calculator(16, 1, n = 45, p = 0.25)




prob_vec <- c(p2, p3, p4, p5, p6, p7)

baseball_test <- chisq.test(x = counts, prob_vec) %>% 
  tidy()
```

Our test statistic is `r round(baseball_test$statistic, 4)`.  

## P-value 
```{r, message=FALSE, warning=FALSE}
pvalue <- 1 - pchisq(baseball_test$statistic, df = length(prob_vec) - 3) 
```

Degrees of Freedom Calculation = N - x - 1 = 6 - 2 - 1 = 3, where N is the number of classes and x is the number of parameters in the binomial distribution.    

The p-value is $1.63176 x 10^{-5}$     

## Conclusion
With a p-value less than 0.05, we can reject the null hypothesis that this sample belongs to a binomially distributed Random Variable with $p = 0.25$ and $n =$ 45.  There does not seem to be evidence pointing to the fact that this is a binomially distributed sample.    

#Dataset #3: Running the Shapiro-Wilk test to see if the beaver1's ```temp``` variable is normally distributed.    
```{r, warning=FALSE, message=FALSE}
beaver <- as.tibble(datasets::beaver1)

beaver
```

$H_{0}$:  The random sample comes from a population with the normal distribution, with unknown mean and standard deviation.     
$H_{1}$:    The distribution function the $X_{i}$'s is nonnormal.    

##Test Statistic     
```{r, warning=FALSE, message=FALSE}
normal_test <- shapiro.test(x = beaver$temp) %>%
  tidy()
```

Our test statistic is `r round(normal_test$statistic, 3)`.  

##P-Value    
Our p-value is `r round(normal_test$p.value, 3)`.    

##Conclusion    
With a p-value less than 0.05, we have enough evidence to reject the null hypothesis that the random sample comes from a population with unknown mean and standard deviation.  There seems to be evidence to support that the distribution function is nonnormal.  

