---
title: "Assignment 6"
subtitle: "Nonparametric Rank Tests"
author: "Kyle Ligon"
date: "`r Sys.Date()`"
output:
  tufte::tufte_html: default
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
  tufte::tufte_book:
    citation_package: natbib
    latex_engine: xelatex
bibliography: skeleton.bib
link-citations: yes
---

```{r setup, include=FALSE}
library(tufte)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)
```

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(broom)
```


# Problem A

Random samples from each of three different types of light bulbs were tested to see how long the light bulbs lasted, with the following results:

Do these results indicate a significant difference between brands? (use Kruskal-Walis test)

```{r, message=FALSE, warning=FALSE}
a <- c(73, 64, 67, 62)
b <- c(84, 80, 81, 77)
c <- c(82, 79, 71, 75)
titles <- c(rep('A', 4), rep('B', 4), rep('C', 4))
vals <- c(a, b, c)
df <- data.frame(titles, vals)

k_test <- kruskal.test(x = df$vals, g = df$titles) %>%
  tidy()
```

## Hypotheses:      

$H_{0}$: All of the 3 population distributions are identical.    
$H_{1}$: At least one of the population tends to yield larger observations than at least one of the other populations.    

## Test Statistic:      

The test statistic is `r round(k_test$statistic, 4)`.  

## P-Value:     

The p-value is `r round(k_test$p.value, 4)`.

## Conclusion:    

With a p-value less than 0.05, there is enough evidence to suggest that we reject the null hypothesis that the populations distributions are identical.  There does seem evidence to support the claim that at least on of the populations produces larger values than at least on of the other distirbutione.  


# Problem B

A blood bank kept a record of the rate of heartbeats for several blood donors.

1. Is the average of the rate of heartbeats the same for men and women?



2. Is the variation among the men significantly greater than the variation among the women?

```{r, message=FALSE, warning=FALSE}
men <- c(58, 65, 74, 74, 76, 79, 82, 86)
women <- c(66, 68, 67, 69, 72, 73, 74, 75, 76)


```

## Hypotheses:      

$H_{0}$: All of the 3 population distributions are identical.    
$H_{1}$: At least one of the population tends to yield larger observations than at least one of the other populations.    

## Test Statistic:      

The test statistic is `r round(k_test$statistic, 4)`.  

## P-Value:     

The p-value is `r round(k_test$p.value, 4)`.

## Conclusion:    

With a p-value less than 0.05, there is enough evidence to suggest that we reject the null hypothesis that the populations distributions are identical.  There does seem evidence to support the claim that at least on of the populations produces larger values than at least on of the other distirbutione.






```{r bib, include=FALSE}
# create a bib file for the R packages used in this document
knitr::write_bib(c('base', 'rmarkdown'), file = 'skeleton.bib')
```
