---
title: "Assignment 5"
author: "Kyle Ligon"
date: "`r Sys.Date()`"
output:
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
  tufte::tufte_html: default
  tufte::tufte_book:
    citation_package: natbib
    latex_engine: xelatex
bibliography: skeleton.bib
link-citations: yes
---

```{r setup, include=FALSE}
library(tufte)
library(tidyverse)
library(broom)
```



# Problem A- Chi Square Test for Differences in Probabilities
```{r, message=FALSE, warning=FALSE}
men <- c(32, 68)
women <- c(26, 74)

taste <- data.frame(men, women, row.names = c('no likey', 'likey'))
taste_test <- chisq.test(x = taste, correct = FALSE) %>%
  tidy()
```

## Hypotheses:         

$H_{0}$: $p_{men}$ = $p_{women}$     

$H_{1}$: $p_{men}$ \neq $p_{women}$

## Test Statistic    
The test statistic of this Chi-Square test is `r round(taste_test$statistic)`

## Critical Region    
We are looking for a $\chi_{0.975, 1, 1}^2$ and $\chi_{0.025, 1, 1}^2$, which is equal to 8.7652.     

## Conclusion    
With the test statistic not greater than or less than the critical region, we cannot reject the Null hypothesis that the probabilities are the same.  There is not enough evidence to suggest that the two probabilities are different.  

# Problem B- Fisher's Exact Test    

```{r, message=FALSE, warning=FALSE}


```

## Hypotheses:         

$H_{0}$: All _c_ populations have the same median.    

$H_{1}$: At least two of the populations have different medians.    

## Test Statistic    
The test statistic of this Median test is 0.8269.  

## Critical Region    
We are looking for a $\chi_{0.95, 2}^2$, which is equal to 5.9941.     

## Conclusion    
With the test statistic greater than the critical region, we can reject the Null hypothesis that the medians are the same.  There is evidence to suggest that at least two medians are different.  

# Problem C- Chi Square Test for Differences in Probabilities    

```{r, message = FALSE, warning=FALSE}
ase <- c(11, 11, 1)
nyse <- c(24, 11, 0)

stocks <- data.frame(ase, nyse, row.names = c('A', 'B', 'C')) %>%
  t()

stock_test <- chisq.test(x = stocks, correct = FALSE)
```

## Hypotheses:         

$H_{0}$: $p_{ASE}$ = $p_{NYSE}$    

$H_{1}$: At least two of the populations have different populations.    

## Test Statistic    
The test statistic of this Median test is 0.8269.  

## Critical Region    
We are looking for a $\chi_{0.95, 2}^2$, which is equal to 5.9941.     

## Conclusion    
With the test statistic greater than the critical region, we can reject the Null hypothesis that the medians are the same.  There is evidence to suggest that at least two medians are different.  


# Problem D- Chi-Square Test    

```{r, message = FALSE, warning=FALSE}


```

## Hypotheses:         

$H_{0}$: All _c_ populations have the same median.    

$H_{1}$: At least two of the populations have different medians.    

## Test Statistic    
The test statistic of this Median test is 0.8269.  

## Critical Region    
We are looking for a $\chi_{0.95, 2}^2$, which is equal to 5.9941.     

## Conclusion    
With the test statistic greater than the critical region, we can reject the Null hypothesis that the medians are the same.  There is evidence to suggest that at least two medians are different.  

# Problem E- Median Test

```{r, message = FALSE, warning=FALSE, results = 'hide'}
#packages needed
library(agricolae)

sampl_1 <- c(35, 42, 42, 30, 15, 31, 29, 29, 17)
sampl_2 <- c(34, 38, 26, 17, 42, 28, 35, 33, 16)
sampl_3 <- c(17, 29, 30, 36, 41, 30, 31, 23, 38)

sample_col <- c(rep('samp_1', length(sampl_1)), 
                rep('samp_2', length(sampl_2)), 
                rep('samp_3', length(sampl_3)))

medians <- c(sampl_1, sampl_2, sampl_3)

med_frame <- data.frame(sample_col = as.factor(sample_col), 
                          medians = medians) %>%
  as.tibble()

med_test <- Median.test(trt = med_frame$sample_col, 
                        y = med_frame$medians)
```

## Hypotheses:         

$H_{0}$: All _c_ populations have the same median.    

$H_{1}$: At least two of the populations have different medians.    

## Test Statistic    
The test statistic of this Median test is 0.8269.  

## Critical Region    
We are looking for a $\chi_{0.95, 2}^2$, which is equal to 5.9941.     

## Conclusion    
With the test statistic greater than the critical region, we can reject the Null hypothesis that the medians are the same.  There is evidence to suggest that at least two medians are different.  